An Adolescent Mini-Neptune in the Kepler Field

### 2021-06-15 paper/ms.tex
minor
update ms

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} considered
the various false positive scenarios, including those of a foreground
eclipsing binary, a hierarchical triple-star binary system, and a
background eclipsing binary.  \citet{morton_false_2016} reported the
latter scenario to be the only possible case, with a (model-dependent)
probability of $\approx10^{-5}$.  While this statistical validation is
reassuring, ``validated" planets have been demonstrated to be false
positives \citep{shporer_three_2017}.  Most failures of the
statistical validation approach arise when attempting to separate the
smallest stars from identically-sized gas giant planets (CITE).  With a size
of $\approx 3.6\,R_\oplus$, Kepler\,1627b does not meet this
concern.  The fact that we detect a low-mass stellar companion in
archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does not change
the conclusion: the companion contributes 1\% of the flux
observed by Kepler, which combined with the observed transit shape
cannot be invoked as an explanation for the observed data
(Section~\ref{app:companionstar}).

Could the planet be on a high obliquity orbit?
% * how do the spectroscopic vsini and the measured v_rot=2πRstar/Prot compare?
%   * Rstar=0.91+-0.05, 2.606 days gives vsini=17.6km/s (+/-0.97km/s)
% * spectroscopic vsini is 18.8 +/- 4.7km/s from Keck HIRES... "20km/s" from
%   TRES.

Kepler\,1627Ab therefore provides an interesting extremum in the ages of the
Kepler planets, and opens multiple avenues for detailed
study.  In order of ascending difficulty, observing its
Doppler transit ($\Delta v_{\rm RM} \approx 20$\ms) with HIRES or the
upcoming KPF spectrograph could yield a measurement of its orbital
obliquity (CITEP,CITEP).  
(A simpler path to the same result might be to determine whether the
$\approx80\,{\rm ppm}$ residual apparent in our transit fit is caused
by gravity darkening.)
A transmission
spectrum focused on either hydrogen or helium absorption could yield
insight into its atmosphere, which is expected to be outflowing at a
rate of {\bf X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  Finally, the
most challenging measurement that may currently be in reach would be a
measurement of its mass, through either detailed timing analyses of
the Kepler data, or through a high-cadence multi-color radial velocity campaign.  The mass would yield constraints on both the planet's
composition (do the mini-Neptunes form with gaseous envelopes?), as
well as its initial entropy (``hot'' or ``cold''? CITE).

@@

L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
transit fitting.
% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
% that encouraged this analysis.
%
L.G.B{.} acknowledges support by the TESS GI Program, programs
G011103 and G022117, through NASA grants 80NSSC19K0386 and
80NSSC19K1728.
%
L.G.B{.} was also supported by a Charlotte Elizabeth Procter
Fellowship from Princeton University.
%
This study was based in part on observations at Cerro Tololo
Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
Association of Universities for Research in Astronomy (AURA) under a
cooperative agreement with the National Science Foundation.
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%

>>>>

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} considered
the various false positive scenarios, including those of a foreground
eclipsing binary, a hierarchical triple-star binary system, and a
background eclipsing binary.  \citet{morton_false_2016} reported the
latter scenario to be the only possible case, with a (model-dependent)
probability of $\approx10^{-5}$.  While this statistical validation is
reassuring, ``validated'' planets have been demonstrated to be false
positives \citep{shporer_three_2017}.  Most failures of the
statistical validation approach arise when attempting to separate the
smallest stars from identically-sized gas giant planets (CITE).  With a size
of $\approx 3.6\,R_\oplus$, Kepler\,1627b does not meet this
concern.  The fact that we detect a low-mass stellar companion in
archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does not change
the conclusion: the companion contributes 1\% of the flux
observed by Kepler, which combined with the observed transit shape
cannot be invoked as an explanation for the observed data
(Section~\ref{app:companionstar}).

% Could the planet be on a high obliquity orbit?
% * how do the spectroscopic vsini and the measured v_rot=2πRstar/Prot compare?
%   * Rstar=0.91+-0.05, 2.606 days gives vsini=17.6km/s (+/-0.97km/s)
% * spectroscopic vsini is 18.8 +/- 4.7km/s from Keck HIRES... "20km/s" from
%   TRES.

Kepler\,1627Ab therefore provides an interesting extremum in the ages of the
Kepler planets, and opens multiple avenues for detailed
study.  Ground-based observations of its Doppler transit ($\Delta
v_{\rm RM} \approx 20$\ms) with HIRES or the upcoming KPF spectrograph
could yield a measurement of its orbital obliquity (CITEP,CITEP).
Similarly, transit spectroscopy aimed at detecting either hydrogen or
helium absorption could yield insight into the planet's atmosphere,
which is expected to be outflowing at a rate of {\bf
X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  A more challenging
measurement would be a measurement of the planet's mass, through
either detailed timing analyses of the Kepler data, or through a
high-cadence multi-color radial velocity campaign.  The mass would
yield constraints on both the planet's composition (``what is the
extent of the initial gaseous envelope for a typical mini-Neptune?''),
as well as its initial entropy (``hot'' or ``cold''? CITE).

The precision of the Kepler data itself may enable detailed analyses
without even needing to acquire more data.
These could include studies transit timing variations, distortion of
the transit signal due to the star's gravity darkening, 
and interactions between the star and planet exhibited by patterns in
the star's flaring.
On the latter point,
Appendix REFERENCE highlights an odd coincidence: 
in the 97.7 days of available 1-minute cadence Kepler observations
(``short-cadence''), 
24 flares were observed to exceed $0.5\%$ in relative flux.
Cumulatively, these flares spanned a total of 0.27 days.
The brightest flare, highlighted in FIGURE CITE, had a flare occur
after it to within 0.2\% of an orbital period.
This seems {\it a priori} unlikely; in fact, 4 of the 24 flare events
have candidate ``successors'' (events that follow to within 2\% of the
planet's orbital period).
Further analysis would be needed to test the hypothesis of whether or
not the flare time
distribution deviated from a Poisson process...
[like: compare Poisson model versus Poisson+periodic mixture?]

FIXME
(A simpler path to the same result might be to determine whether the
$\approx80\,{\rm ppm}$ residual apparent in our transit fit is caused
by gravity darkening.)

@@

L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
transit fitting, and to R{.}~Kerr for kindly providing us with the
\citet{Kerr2021} membership list prior to its publication.
% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
% that encouraged this analysis.
%
L.G.B{.} acknowledges support by the TESS GI Program, programs
G011103 and G022117, through NASA grants 80NSSC19K0386 and
80NSSC19K1728.
%
L.G.B{.} was also supported by a Charlotte Elizabeth Procter
Fellowship from Princeton University.
%
This study was based in part on observations at Cerro Tololo
Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
Association of Universities for Research in Astronomy (AURA) under a
cooperative agreement with the National Science Foundation.
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%
